Analysis of epistatic inflation under large single additive effects, with a focus on the replication rate

### 2019-09-30 null_replication/scripts/analyse_results.rmd
updated analysis of null simulations

@@

```{r}
library(tidyverse)
load("../data/rs67903230.rdata")
```

@@

```{r}
plot(lambda_disc ~ varexp, res)
```

@@

```{r}
hist(res$nsig_disc_bonf)
```

@@

@@

```{r}
plot(nsig_disc_bonf ~ varexp, res)
```

@@

@@

We can attempt to replicate at a Bonferroni threshold (within one simulation, how many associations were attempted to replicate at corrected p < 0.05) or FDR (within one simulation, how many associations repliace at FDR < 0.05).

Note that these are much more relaxed thresholds than what we used in the original paper - which was 0.05 / 501.

```{r}

res$lambda_disc_bin <- cut(res$lambda_disc, 5)
temp <- res %>%
	group_by(lambda_disc_bin, nsig_disc_bonf) %>%
	summarise(count=n(), rate=n()/sum(res$lambda_disc_bin == .$lambda_disc_bin[1]))

ggplot(temp %>% filter(nsig_disc_bonf > 0), aes(x=lambda_disc_bin, y=count)) +
geom_bar(stat="identity", aes(fill=nsig_disc_bonf))

table(res$lambda_disc_bin)
hist(res$lambda_disc)

plot(cond_bonf_bonf2 ~ varexp, res)

plot(I(cond_bonf_bonf2/nsig_disc_bonf) ~ varexp, res)

mean(res$cond_bonf_bonf / res$nsig_disc_bonf, na.rm=T)
mean(res$cond_bonf_bonf2 / res$nsig_disc_bonf, na.rm=T)

res$bonf_bonf_rr <- res$cond_bonf_bonf / res$nsig_disc_bonf
res$bonf_fdr_rr <- res$cond_bonf_fdr / res$nsig_disc_bonf
res$bonf_bonf2_rr <- res$cond_bonf_bonf2 / res$nsig_disc_bonf

mean(res$bonf_bonf_rr, na.rm=TRUE)
mean(res$bonf_bonf2_rr, na.rm=TRUE)
tapply(res$bonf_bonf_rr, res$lambda_disc_bin, function(x) mean(x, na.rm=T))
tapply(res$bonf_fdr_rr, res$lambda_disc_bin, function(x) mean(x, na.rm=T))

tapply(res$bonf_bonf_rr, res$nsig_disc_bonf, function(x) mean(x, na.rm=T))
tapply(res$bonf_fdr_rr, res$nsig_disc_bonf, function(x) mean(x, na.rm=T))

```

>>>>

## Summary

These simulations suggest

- that discovering associations at experiment wide thresholds (e.g. 1e-16) is unlikely even when the lambda is very high
- the replication rate could be quite high using FDR
- the replication rate is very low at experiment wide thresholds
- we might get higher experiment wide replication rates if we found a way that the discovery effects surpassed the experiment wide threshold

@@

We also added an extra variable to test. Most expression traits have a large cis effect and a complex, polygenic trans component. For half of the simulations we also created a polygenic component that explained 30% of the variance, which was composed of 1000 SNPs which each had an effect sampled from a normal distribution.

```{r}
library(tidyverse)
load("../data/rs67903230.rdata")
res$polygenic <- res$i %% 2 == 0
```

@@

```{r}
ggplot(res, aes(x=varexp, y=lambda_disc)) +
geom_point(aes(colour=polygenic)) +
geom_smooth(aes(colour=polygenic))
```

@@

```{r}
ggplot(res, aes(x=nsig_disc_bonf)) +
geom_histogram(aes(fill=polygenic), position="dodge")
```

@@

There looks to be no affect of having a polygenic trans component, so from now on we will not stratify the results on that basis.

@@

```{r}
ggplot(res, aes(x=varexp, y=nsig_disc_bonf)) +
geom_point()
```

@@

Here's a look at the average number of discovery hits per simulation, binned by levels of genome wide lambda

```{r}
res$lambda_disc_bin <- Hmisc::cut2(res$lambda_disc, g=30)
group_by(res, lambda_disc_bin) %>% summarise(n=mean(nsig_disc_bonf, na.rm=T)) %>% 
ggplot(aes(x=lambda_disc_bin, y=n)) +
geom_point() +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Simulation's lambda", y="Mean number of discovery hits")
```

@@

We can attempt to replicate at

* Bonferroni threshold (within one simulation, how many discovery associations were replicated at Bonferroni corrected p < 0.05)
* FDR (within one simulation, how many discovery associations replicate at FDR < 0.05).
* experiment threshold (within one simulation, how many discovery associations replicated at p < 0.05/501, the threshold used in the paper)

```{r}
res$bonf_bonf_rr <- res$cond_bonf_bonf / res$nsig_disc_bonf
res$bonf_fdr_rr <- res$cond_bonf_fdr / res$nsig_disc_bonf
res$bonf_bonf2_rr <- res$cond_bonf_bonf2 / res$nsig_disc_bonf

repl_rate <- bind_rows(
	group_by(res, lambda_disc_bin) %>% summarise(m=mean(bonf_bonf_rr, na.rm=T)) %>% mutate(threshold="Bonferroni"),
	group_by(res, lambda_disc_bin) %>% summarise(m=mean(bonf_fdr_rr, na.rm=T)) %>% mutate(threshold="FDR"),
	group_by(res, lambda_disc_bin) %>% summarise(m=mean(bonf_bonf2_rr, na.rm=T)) %>% mutate(threshold="Experiment")
)
repl_rate %>% ggplot(aes(x=lambda_disc_bin, y=m, group=threshold)) +
geom_line(aes(colour=threshold)) +
scale_colour_brewer(type="qual") +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Simulation's lambda", y="Replication rate")
```

Overall, using FDR there is a reasonably high replication rate which doesn't correspond to discovery lambda. But the experiment wide threshold can only be met when the discovery lambda is very high, but even then the replication rate is extremely low.
